---
title: "writeup"
author: "Chaeyeong"
date: "2024-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Design
Research Question
How is financial inclusion (e.g., account ownership, saving, borrowing, and digital payments) related to poverty levels across countries with complete data for 2014, 2017, and 2021?
Research Objectives
To examine the correlation between financial inclusion variables and poverty levels.
To control for economic factors like GDP and assess their moderating effect on poverty.
To evaluate temporal changes in the relationship between financial inclusion and poverty over the three years.
Key Variables
Dependent Variable:
Poverty Level: poverty_headcount_ratio
Independent Variables (Financial Inclusion):
per_account_ownership_poor: Percentage of poor individuals with a financial account.
per_saving_poor: Percentage of poor individuals saving in financial institutions.
per_borrowing_poor: Percentage of poor individuals borrowing from financial institutions.
per_digital_payment_poor: Percentage of poor individuals making digital payments.
Control Variables:
Economic Growth: gdp (log-transformed for better interpretation).
Year Effects: Dummy variables for 2014, 2017, and 2021 to control for time trends.
Empirical Strategy
Model Specification:
Use a fixed-effects (FE) panel regression model to control for unobserved, time-invariant country-specific characteristics (e.g., geography, culture).
Where:


poverty_headcount_ratioit = Î²0 +Î²1(per_account_ownership_poorit)+Î²2 (per_saving_poori)+ Î²3(per_borrowing_poorit)+Î²4(per_digital_payment_poorit)+
Î²5(gdpit)+Î±i+Î³t+Ïµit
i: Country.
t: Year.
ğ›¼ğ‘–: Country fixed effects.
ğ›¾ğ‘¡: Year fixed effects.
ğœ–ğ‘–ğ‘¡: Error term

1. Balanced Panel Data Analysis
Focus your analysis on the 47 countries with data for all three years.
This design enables you to apply fixed-effects or random-effects models for a balanced panel, which provides robust insights into the temporal relationships between financial inclusion variables and poverty across years.
Advantages:
Robust temporal and cross-sectional comparisons.
Controls for country-level unobserved heterogeneity.
Disadvantages:
Reduces the sample size, potentially limiting the generalizability of results.
Implementation:

 

```{r}

```


